---
title: "Project 5 - Title"
author: "Team 13"
output:
  html_document: default
  html_notebook: default
---

### Project Description
  
  
### Why did we choose this project?
  
  
### Dataset Description
Those data have been recorded in Brazil, in the state of Espirito Santo. Datapoints are only from the public sector, primary care.  

*Description of the dataset*   
- Age: Age of the patient  
- Gender: Gender of the patient  
- AppointmentRegistration: Time and date when the patient took the appointment  
- AppointmentData: Date of the appointment  
- Diabetes: Is the patient affected by diabetes? (0/1)  
- Alcoolism: is the patient alcoolic? (0/1)  
- Hipertension: Is the patient affected by hypertension? (0/1)  
- Handicap: Level of handicap of the aptient (0/1/2/3/4)  
- Smokes: Is the patient smoking? (0/1)    
- Scholarship: Is the patient receiving a scholarship? Those scholarships are given by Bolsa Familia to low income families who accept to send their kids to school nd have them vaccinated.   
- Tuberculosis: Is the patient affected by tuberculosis? (0/1)  
- AwaitingTime: How many days the patient waited between the appointment registration and the date of the appointment  
- Status: Did the patient show-up or not? (No-Show/Show-Up)  
  
*Source*    
It's a kaggle dataset which have been taken here. https://www.kaggle.com/joniarroba/noshowappointments

### Data Cleaning

**Step 0: Load Packages**

```{r, message=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(gridExtra)
library(caret)
```

**Step 1: Load Dataset**
```{r}
# Medical Dataset
dataset <- read.csv("../data/No-show-Issue-Comma-300k.csv")
```

**Step 2: Clean & Organize Dataset**
```{r}
# Check Format
str(dataset)
```

```{r}
# Update Columns Names
colnames(dataset) <- c("age", "gender", "appointment_registration", 
                       "appointment_date", "dayofweek_apptmt", "show_up",
                       "diabetes", "alcoholism", "hypertension", "handicap", 
                       "smoker", "scholarship", "tuberculosis",
                       "sms_reminder", "daydiff_regist_appt")

# Update Columns Format
dataset$sms_reminder <- factor(dataset$sms_reminder)
dataset$diabetes <- factor(dataset$diabetes)
dataset$alcoholism <- factor(dataset$alcoholism)
dataset$hypertension <- factor(dataset$hypertension)
dataset$handicap <- factor(dataset$handicap)
dataset$smoker <- factor(dataset$smoker)
dataset$scholarship <- factor(dataset$scholarship)
dataset$tuberculosis <- factor(dataset$tuberculosis)
```

```{r}
# Check Values
summary(dataset)
```

```{r}
# Data Cleaning

# Age: The Age shouldn't be negative, we will remove those rows. 
dataset <- dataset[dataset$age>0,]

# Appointment day of the week: We don't know if this variable is linked to the registration date or the apointment date - we will delete it and use the raw columns only
dataset$dayofweek_apptmt <- NULL

# Number of days between registration time and appointment date: We will convert this value to be positive
dataset$daydiff_regist_appt <- -(dataset$daydiff_regist_appt)

# Show-up / No-Show: 
dataset$show_up <- revalue(dataset$show_up, c("No-Show"=0, "Show-Up"=1))

# Clean date and time:
## Appt Registration Date
dataset$appointment_registration <- as.POSIXct(as.character(dataset$appointment_registration), 
                                               "%Y-%m-%d T%H:%M:%SZ", 
                                               tz ="Brazil/East")

# Set language to English 
Sys.setlocale("LC_TIME", "C")

dataset$weekday_registration <- as.factor(format(dataset$appointment_registration, "%a"))
dataset$day_registration <- as.numeric(as.integer(format(dataset$appointment_registration, 
                                                         "%d")))
dataset$month_registration <- as.numeric(format(dataset$appointment_registration, "%m"))
dataset$hour_registration <- as.numeric(format(dataset$appointment_registration,"%H"))
dataset$minute_registration <- as.numeric(format(dataset$appointment_registration,"%M"))
dataset$day_of_year_registration <- as.numeric(format(dataset$appointment_registration,"%j"))
dataset$week_of_year_registration <- as.numeric(format(dataset$appointment_registration,"%U"))
dataset$am_pm_registration <- as.factor(format(dataset$appointment_registration,"%p"))

## Appt Actual Date
dataset$appointment_date <- as.POSIXct(as.character(dataset$appointment_date), 
                                       "%Y-%m-%d T%H:%M:%SZ", 
                                       tz = "Brazil/East")

dataset$weekday_appointment <- as.factor(format(dataset$appointment_date,"%a"))
dataset$day_appointment <- as.numeric(as.integer(format(dataset$appointment_date,"%d")))
dataset$month_appointment <- as.numeric(format(dataset$appointment_date,"%m"))
dataset$day_of_year_appointment <- as.numeric(format(dataset$appointment_date,"%j"))
dataset$week_of_year_appointment <- as.numeric(format(dataset$appointment_date,"%U"))
```

### Exploratory Data Analysis

**Number of Datapoints**
```{r}
# Number of rows
nb_rows <- nrow(dataset)
```  
We have access to 300k datapoints.  

**Prediction Variable**  
Let's first analyse the prediction variable alone.    
```{r}
ggplot(dataset, aes(x = show_up, fill=show_up)) +
  geom_bar() +
  ggtitle("Number of Patients Showing Up") +
  ylab("Count") +
  xlab("Show-up") + 
  scale_fill_discrete("Show-up", labels=c("No", "Yes")) +
  scale_x_discrete(labels=c("No", "Yes"))
```  

On the 300k datapoints we have, one third of them are No-Show Datapoints.   
From a data science point of view, we have unbalanced classes, we will have to deal with it when doing predictions.   

**Timeframe & Number of Datapoints per Day**   
```{r}
dataset %>% 
  group_by(appointment_date) %>% 
  summarise(patients=n()) %>% 
  ggplot(aes(x=appointment_date, y=patients)) + 
    geom_point(alpha=0.3) +
    xlab("Appointment Date") +
    ylab("Number of Patients") +
    ggtitle("Number of Patients per Date") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```   

We have access to 2 years of data from January 2014 or December 2015. THe number of datapoints per date is around constant. Outliers are Sundays.  

**Gender Analysis**   
```{r}
gender_number <- ggplot(dataset, aes(x=gender)) +
  geom_bar() +
  xlab("Gender") + 
  ylab("Count")

gender_ratio <- ggplot(dataset, aes(x=gender, fill=show_up)) + 
  geom_bar(position="fill") + 
  xlab("Gender") + 
  ylab("Ratio") +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))

grid.arrange(gender_number, gender_ratio, ncol=2, top="Gender Analysis")
```    

The first insight is that women are more often going to visit a physician. However the ration of no-show/show are realyy similar. Gender doesn't seem to play an important role.  

**Age Analysis**   
```{r}
dataset %>% 
  group_by(age) %>% 
  summarise(nb_perage=n()) %>% 
  ggplot(aes(x=age, y=nb_perage)) + 
    geom_point(alpha=0.3) +
    xlab("Age") +
    ylab("Number of Patients") +
    ggtitle("Number of Patients depending on the Age")
```     

On the first graph, we display the number of patient depending on the age to know what the sample we have look like. After 60 years old, people begin to die, and hence there are less patients. We see on the representation of the ife expectancy of people in Brazil whicih is comprised between 60 and 100 years old. (The life expectancy in Brazil is 73.8 years old).   
   
```{r}
ggplot(dataset, aes(x=show_up, y=age, fill=show_up)) + 
  geom_boxplot() +
  xlab("Show-up") +
  ylab("Age") +
  ggtitle("Proportion of people showing up depending on the Age") +
  scale_fill_discrete("Show-up", labels=c("No", "Yes")) +
  scale_x_discrete(labels=c("No", "Yes"))
```    

The age seems to be a important factor. The older the people are the more likely they are going to show up.    
   
```{r}
by_age <-
  dataset %>%
    group_by(age) %>%
    summarise(show = sum(show_up==1), noshow = sum(show_up == 0))
by_age$ratio = by_age$noshow / (by_age$show+by_age$noshow)

ggplot(data = by_age, aes(x = age, y = ratio)) + 
  geom_line() + 
  geom_smooth(method = "glm")
```    

Above, we can see the same trend. After 95 years old the figures are not significant because the sample is too small. Hence only the first part of the graph can be interpreted.    
   
**Age & Gender Analysis**   
```{r}
ggplot(data = dataset, aes(x = age, fill = show_up)) + 
  geom_bar() + 
  facet_wrap(~gender) +
  xlab("Age") +
  ylab("Count") +
  ggtitle("Number of Patients showing up depending on Age and Gender") +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
```     

The ratio is similar between woman and man. We see again that women take more appointments. However we see that before 20 years old the numbe rof appointments and the trend is exactly the same.    
   
**Health Conditions Analysis**   
```{r}
diabetes <- ggplot(dataset, aes(x=diabetes, fill=show_up)) + 
  geom_bar(position="fill") +
  ylab("Ratio") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
alcoholism <- ggplot(dataset, aes(x=alcoholism, fill=show_up)) + 
  geom_bar(position="fill") +
  ylab("Ratio") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
hypertension <- ggplot(dataset, aes(x=hypertension, fill=show_up)) + 
  geom_bar(position="fill") +
  ylab("Ratio") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
handicap <- ggplot(dataset, aes(x=handicap, fill=show_up)) + 
  geom_bar(position="fill") +
  ylab("Ratio") +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
smoker <- ggplot(dataset, aes(x=smoker, fill=show_up)) + 
  geom_bar(position="fill") +
  ylab("Ratio") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
tuberculosis <- ggplot(dataset, aes(x=tuberculosis, fill=show_up)) + 
  geom_bar(position="fill") +
  ylab("Ratio") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))

grid.arrange(diabetes, tuberculosis, hypertension, handicap, smoker,
             alcoholism, ncol=2, top='Health Conditions')
```      

Form this first anaysis, we can conclude that people having a handicap are more likely to show up, going further the more severe is there handicap the more likely they are going to show up. The same trand is observed for people havnig diabete or hypertension.  
However for smokers, alcoholic people and people suffering from tuberculosis, the trend is reverse.  
To validate this insights, we need to look at the samples size, which we are going to do below.  
   
```{r}
diabetes <- ggplot(dataset, aes(x=diabetes, fill=show_up)) + 
  geom_bar() +
  ylab("Count") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
alcoholism <- ggplot(dataset, aes(x=alcoholism, fill=show_up)) + 
  geom_bar() +
  ylab("Count") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
hypertension <- ggplot(dataset, aes(x=hypertension, fill=show_up)) + 
  geom_bar() +
  ylab("Count") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
handicap <- ggplot(dataset, aes(x=handicap, fill=show_up)) + 
  geom_bar() +
  ylab("Count") +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
smoker <- ggplot(dataset, aes(x=smoker, fill=show_up)) + 
  geom_bar() +
  ylab("Count") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))
tuberculosis <- ggplot(dataset, aes(x=tuberculosis, fill=show_up)) + 
  geom_bar() +
  ylab("Count") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))

grid.arrange(diabetes, tuberculosis, hypertension, handicap, smoker,
             alcoholism, ncol=2, top='Health Conditions')
```     

As we were expected, we always have less sick people. The number of people having tuberculosis is so low that the ratio we analyzed earlier should not to be taken into account.   
   
**SMS Reminder**    
```{r}
smsreminder_number <- ggplot(dataset, aes(x=sms_reminder)) +
  geom_bar() +
  xlab("Sms Reminder") + 
  ylab("Count")

smsreminder_ratio <- ggplot(dataset, aes(x=sms_reminder, fill=show_up)) + 
  geom_bar(position="fill") + 
  xlab("Sms Reminder") + 
  ylab("Ratio") +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))

grid.arrange(smsreminder_number, smsreminder_ratio, ncol=2, top="Sms Reminder")
```     

There is no strong pattern associated with the sending of a sms reminder. Without taking into account people that received 2 sms (too few people), we notice that the ration is really similar. However we don't know if those sms are sent randomnly or not.   
   
**Scholarship Analysis**   
```{r}
smsreminder_number <- ggplot(dataset, aes(x=scholarship)) +
  geom_bar() +
  xlab("Scholarship") + 
  ylab("Count")

smsreminder_ratio <- ggplot(dataset, aes(x=scholarship, fill=show_up)) + 
  geom_bar(position="fill") + 
  xlab("Scholarship") + 
  ylab("Ratio") +
  scale_fill_discrete("Show-up", labels=c("No", "Yes"))

grid.arrange(smsreminder_number, smsreminder_ratio, ncol=2, top="Scholarship")
```     
  
We did explain at the beginning of this file what scholarship meant, it's basically linked to the level of income of the family. We see that this feature have a strong impact.  

**Time Analysis**

Overview of the dataset per date.  
```{r}
dataset$year_appointment <- as.numeric(format(dataset$appointment_date,"%Y"))

dataset %>% 
  group_by(day_of_year_appointment, year_appointment) %>% 
  summarise(patients=n()) %>% 
  ggplot(aes(x=day_of_year_appointment, y=patients)) + 
    geom_point(alpha=0.3) +
    geom_smooth(method = 'loess') +
    facet_grid(year_appointment ~ .) +
    xlab("Day of the Year") +
    ylab("Number of Appointments") +
    ggtitle("Number of Appointments per Day")
```

We observe a very small trend: people schedule slightly more medical appointment during the first half of the year. 

Now, to explain a little bit how time influences our variable of interest, we show the evolution of the noshow ratio through the period we are studying.

```{r}
dataset %>% 
  group_by(day_of_year_appointment, year_appointment) %>% 
  summarise(total_noshow=sum(show_up==0)/n()) %>% 
  ggplot(aes(x=day_of_year_appointment, y=total_noshow)) + 
    geom_line(alpha=0.5, color = "red") +
    geom_smooth(method = 'loess', se = FALSE, color = alpha('black',0.5)) +
    facet_grid(year_appointment ~ .) +
    xlab("Day of the Year") +
    ylim(0,1) +
    ylab("Percentage of No-Show") +
    ggtitle("Percentage of No-Show per Day")
```

Using a daily granularity, we observe a high variance that makes the graph difficult to interpret. Let's change modify the graph by using a weekly granularity.  

```{r}
dataset %>% 
  group_by(week_of_year_appointment, year_appointment) %>% 
  summarise(total_noshow=sum(show_up==0)/n()) %>% 
  ggplot(aes(x=week_of_year_appointment, y=total_noshow)) + 
    geom_line(alpha=0.5, color = "red") +
    geom_smooth(method = 'loess', se = FALSE, color = alpha('black',0.5)) +
    facet_grid(year_appointment ~ .) +
    xlab("Day of the Year") + 
    ylim(0,1) +
    ylab("Percentage of No-Show") +
    ggtitle("Percentage of No-Show per Week")
```

On a weekly basis we have more or less the same observations: we don't see any correlation between the date and the no-show ratio having a year window. 

The natural next step is to look at the same ratio vs weekdays. It sounds pretty intuitive that there are some information to extract.
```{r}
by_day_of_week = group_by(dataset, weekday_appointment)
by_day = summarise(by_day_of_week, 
                   show = sum(show_up == 1), 
                   noshow = sum(show_up == 0))
by_day$ratio = by_day$noshow / (by_day$show+ by_day$noshow)
by_day$weekday_appointment = relevel(by_day$weekday_appointment, "Sun")
by_day$weekday_appointment = relevel(by_day$weekday_appointment, "Sat")
by_day$weekday_appointment = relevel(by_day$weekday_appointment, "Fri")
by_day$weekday_appointment = relevel(by_day$weekday_appointment, "Thu")
by_day$weekday_appointment = relevel(by_day$weekday_appointment, "Wed")
by_day$weekday_appointment = relevel(by_day$weekday_appointment, "Tue")
by_day$weekday_appointment = relevel(by_day$weekday_appointment, "Mon")

qplot(weekday_appointment, ratio, data = by_day, 
      geom = c("boxplot"), 
      xlab = "Day of the Week",
      ylab = "Percentage of No-Show",
      main = "Percentage of No-Show depending on the Day")

```

We clearly see a trend here: the ratio is extremely high on Saturday! It jumps from 30 to 37%. Sunday is not significant because very few people have appointment on Sundays.    

Let us have a look ot the moment in the day the patient register for the appointment.
```{r}
by_hour <- group_by(dataset, hour_registration)
by_hour <- summarise(by_hour, 
                     show=sum(show_up==1), 
                     noshow=sum(show_up == 0))
by_hour$ratio <- by_hour$noshow / (by_hour$show+ by_hour$noshow)

ggplot(data=by_hour[by_hour$hour_registration>5,][by_hour$hour_registration<21,],
       aes(x = hour_registration, y = ratio)) + 
  geom_line() + 
  ylim(0,0.5) + 
  xlab("Hour (24h convention)") + 
  ylab("Percentage of No-Show") +
  ggtitle("Percentage of No-Show depending on the Time of the Day")
```
We notice that people calling very early (<6am) in the morning are less likely to not show up compare to people calling late at night (>7pm). 

Another important aspect in the time dimention is the time difference between the moment they register for the appointment and the appointment date. 
```{r}
by_diff <- group_by(dataset, daydiff_regist_appt)
by_diff <- summarise(by_diff, 
                     show=sum(as.numeric(show_up)), 
                     noshow=sum(show_up == 0))
by_diff$ratio <- by_diff$noshow / by_diff$show

ggplot(data = by_diff, aes(x = daydiff_regist_appt, y = ratio)) + 
  geom_line() + 
  xlab("Day Difference") + 
  ylab("Percentage of No-Show") + 
  ggtitle("Percentage of No-Show depending on the number of Waiting Days")

```

We first notice that the curve is extremely heratic after a certain value due to the number of datapoints. Let us zoom in the 0-120 days window. A waiting time of more than 120 days is really rare, so the second part od the curve isn't significant.  
```{r}
by_diff <- group_by(dataset, daydiff_regist_appt)
by_diff <- summarise(by_diff, 
                     show = sum(show_up==1), 
                     noshow = sum(show_up == 0))
by_diff$ratio <- by_diff$noshow / (by_diff$show+by_diff$noshow)
by_diff$daydiff_regist_appt_squared <- by_diff$daydiff_regist_appt**2

ggplot(data = by_diff[by_diff$daydiff_regist_appt<120,], 
       aes(x = daydiff_regist_appt, y = ratio)) + 
  geom_line() + 
  geom_smooth(method = 'loess') +
  xlab("Day Difference") + 
  ylab("Percentage of No-Show") + 
  ggtitle("Percentage of No-Show depending on the number of Waiting Days")

```

It seems that we have a good model using a polynomial regression as follow:

```{r}
summary(lm(ratio~daydiff_regist_appt + daydiff_regist_appt_squared, by_diff[by_diff$daydiff_regist_appt<120,]))
```

We see that the polynomial model fit pretty well this part of the curve with 2 p-values extremely low, which is good.  

**Weather Analysis**

In this part, we want to assess the impact of the weather on attendance of patients. We webscraped data about the weather in this specific city during the period of the analysis; and we look for any correlation between the level of precipitation, the temperature, the visibility, the wind speed and the noshow ratio previously described.  
The webscrabing code can be foudn in the lib folder (Weather scrapping.ipynb).  

```{r}
# Load Weather data
load("../output/with_weather_dataset.RData")
```

*Precipitation level*
```{r}
by_temp <- group_by(dataset, appointment_date)
by_temp <- summarise(by_temp, 
                     show = sum(show_up==1), 
                     noshow = sum(show_up == 0), 
                     max_temp = max(max_temp_faren), 
                     lowest_visibility = max(as.numeric(lowest_visibility_miles)), 
                     lowest_visibility = max(as.numeric(lowest_visibility_miles)),
                     wind = max(max_wind_mph), 
                     max_precipitation_inches = max(precipitation_inches))
by_temp$ratio <- by_temp$noshow / (by_temp$show + by_temp$noshow)

qplot( max_precipitation_inches, ratio, 
       data = by_temp[by_temp$max_precipitation_inches<2,],
       ylab = "Percentage of No-Show",
       xlab = "Precipitation (in inches)") + 
  geom_smooth(show.legend = TRUE,method = 'lm')

summary(lm(ratio~max_precipitation_inches, 
           data = by_temp[by_temp$max_precipitation_inches<2,]))
```

Each point on the graph represent one day. We note that there is no correlation between the ratio and the precipitation level.  

*Temperature*
```{r}
qplot(max_temp, 
      ratio, 
      data=by_temp,
      ylab="Temperature",
      xlab="Percentage of No-Show") + 
  geom_smooth(show.legend=TRUE, 
              method='lm')

summary(lm(ratio ~ max_temp, data = by_temp))
```

Same result with the temperature.

*Visibility*
```{r}
qplot(lowest_visibility, 
      ratio, 
      data=by_temp,
      xlab="Lowest Visibility of the Day",
      ylab="Percentage of No-Show") + 
  geom_smooth(show.legend=TRUE, 
              method='lm')

summary(lm(ratio ~ lowest_visibility, data = by_temp))
```

Visibility IS correlated with the ratio. We observe that the ratio decrease when the lowest visibility of the day increase. We obtain a good p-value of 3% for this correlation.

*Wind*
```{r}
qplot(wind, 
      ratio, 
      data=by_temp) + 
  geom_smooth(show.legend=TRUE, 
              method='lm')

summary(lm(ratio ~ wind, data = by_temp))
```

Just like the two first variables, the wind speed is not correlated with the No-Sow ratio. 

We conclude from this analysis that among all weather variables we found, the only feature that is a little bit correlated with the ratio is the lowest visibility of the day.  